Revisited hidden weight bit function

### 2024-11-16 RevisedHiddenWeight.tex
Updates from Overleaf
some part 6, bib, and pre-intro

@@

When $n$ is even, the nonlinearity of a function can reach at most $2^{n-1}-2^{n/2 -1}$. Functions reaching this maximum are called bent functions. The quadratic function $d_n$ given by: $d_n(x)=\sum_{i \text{ odd}}^n x_i x_{i+1}$ is an example of bent function.

@@

@@

\begin{theorem}[Balancedness of $f$]
	Let $n \in \N^*$ even, and $f$ be the $n$-variable function from Definition~\ref{def:revHWBF}, the following holds on its Walsh transform in $0_n$:
	\[\wt{f}(0_n)= \left \{
	\begin{array}{l l}
	0 & \text{ if } n/2 \text{ is even, } \\
	-2 \binom{n/2-1}{(n-2)/4} & \text{ if }  n/2 \text{ is odd.}
	\end{array}\right.\]

@@

\section{Experiments and comparisons}

@@

\pmnote{under construction}

@@

\pmnote{under construction, waiting for the parameters for the FAI}

\subsection{Degree}

@@

@@

\subsection{Algebraic immunity}

@@

From~\cite{DAM:WCST14} Theorem 4, the algebraic immunity of the HWBF is at least $\lfloor n/3\rfloor +1$, which leads to $\lfloor n/3\rfloor -1$ for $f$.

@@

>>>>

recent developpement stream ciphers

\cite{INDO:HofMeaRic20,CCS:CDPP22,CIC:MeaParPei24}

HHE \cite{CCS:NaeLauVai11}

FLIP~\cite{EC:MJSC16}

RASTA~\cite{C:Rasta}

FiLIP~\cite{INDO:MCJS19}

HWBF:

\cite{IEEE:Bryant91}

\cite{IEEE:Carlet22}

\cite{eprint:CarPar23}

\cite{DAM:MeaOza24}

@@

When $n$ is even, the nonlinearity of a function can reach at most $2^{n-1}-2^{n/2 -1}$. Functions reaching this maximum are called bent functions, due to their multiple applications and importance they have been the focus of multiple works \eg \cite{Rothaus76,FSE:Dobbertin95,DBLP:conf/itw/BudaghyanCH11,Tokareva15,Mesnager16}. 
The quadratic function $d_n$ given by: $d_n(x)=\sum_{i \text{ odd}}^n x_i x_{i+1}$ is an example of bent function.

The algebraic degree of multiples of a Boolean function is a crucial criterion for various attacks on stream ciphers that utilize a Boolean function as a filter. The well-known algebraic attack~\cite{EC:CouMei03} and fast algebraic attack~\cite{C:Courtois03} on filtered LFSRs have motivated the study of cryptographic properties such as algebraic immunity and fast algebraic immunity, investigated in numerous works (\eg \cite{C:HawRos04,ISC:JiaZhaWan13})).

Nowadays, these parameters are systematically determined for any function considered as a filter.

@@

\begin{definition}[Fast algebraic immunity, \eg \cite{EC:ACGKMR06}]\label{def:fai} The fast algebraic immunity of a Boolean function $f\in \Bfn{n}$ is defined as $\FAI(f)=\min \{ 2 \AI(f), \min _{1 \le \degg(g)<\AI(f)} (\degg(g)+ \degg(fg))\}$.
\end{definition}

@@

\begin{theorem}[Balancedness of $f$]\label{th:bal}
	Let $n \in \N^*$ even, and $f$ be the $n$-variable function from Definition~\ref{def:revHWBF}, the following holds on its Walsh transform in $0_n$:
	\[\wt{f}(0_n)= \left \{
	\begin{array}{l l}
	0 & \text{ if } n/2 \text{ is even, } \\
	-2 \binom{n/2-1}{(n-2)/4} & \text{ if }  n/2 \text{ is odd.}
	\end{array}\right.\]

@@

\section{Experiments and comparisons}\label{sec:expAndComp}

@@

@@

In this part we give the other cryptographic parameters of the function $f$. Since its balancedness is characterized by Theorem~\ref{th:bal} and its nonlinearity deeply analyzed in Section~\ref{sec:Cauchy} and Section~\ref{sec:expAndComp}, in the following we focus on the algebraic properties. More precisely we analyze the degree, algebraic immunity and fast algebraic immunity of $f$, providing comparisons with other weightwise quadratic functions.

%\pmnote{under construction, waiting for the parameters for the FAI}

%\subsection{Degree}

@@

In Table~\ref{table:comparisonsDeg}, we display the algebraic degree of different weightwise quadratic function. We observe that $f$ is the function with higher degree, jointly with the HWBF.

@@

%\subsection{Algebraic immunity}

@@

From~\cite{DAM:WCST14} Theorem 4, the algebraic immunity of the HWBF is at least $\lfloor n/3\rfloor +1$, which leads to $\lfloor n/3\rfloor -1$ for $f$. 
In Table~\ref{table:comparisonsAI}, we present the algebraic immunity of various weightwise quadratic functions. We observe that the function $f$ demonstrates the best performance after the majority function, which is known to achieve optimal algebraic immunity.

@@

Additionally, we considered the fast algebraic immunity of $f$ (see Definition~\ref{def:FAI}). In Table~\ref{table:FAI} we give the best couples $(d,e)$ encountered for the function $f$, where $fg=h$, $\degg(g)=d, \degg(h)=e$.

These various results on the algebraic properties of $f$ suggest that it possesses strong resistance against standard attacks. Additionally, its high nonlinearity makes it a well-suited candidate for use as a filter function in contexts such as filtered LFSRs or in homomorphically-friendly schemes like FLIP and FiLIP.

\begin{table}[h]
	\centering
	\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
		\hline
		$ n $      & $4$     & $6$     & $8$     & $10$    & $12$    & $14$    & $16$    & $18$    \\ \hline
		$(d, e)$ & $(1,2)$ & $(1,3)$ & $(1,5)$  & $(1,7)$ & $(1,9)$ & $(1,11)$ & $(1,13)$ & $(1,15)$ \\ 
		 &     &    & $(2,4)$  & $(2,6)$ & $(2,8)$ & $(2,10)$ & $(2,12)$ & $(2,14)$\\ 
		 &     &    &     & $(3,5)$ & $(3,6)$ & $(3,9)$ & $(3,11)$ & $(3,13)$\\ 
		 &     &    &     &  &  &  & $(4,9)$ & $(4,11)$\\ 
	 &     &    &     &  &  &  &  & $(5,10)$ \\ \hline
	\end{tabular}
	\caption{Lowest possible values of $(d, e)$ for different values of $ n$}
	\label{table:FAI}
\end{table}
